Assessing the impact of urban heating on tree growth in Berlin with open inventory and environmental data

### 2022-05-29 analysis/paper/paper.Rmd
updating paper

<!-- <!-- lacked spatially-explicit effect size estimates across the urban-rural space and  --> -->

@@

It features nearly 650000 individuals covering 94 genera and some 600 species and/or cultivars, listing information on location, stem diameter (at breast height; $DBH$), and stem height, among other variables, for the majority of street and park trees.

@@

caption <- "Data description used for maps/visualizations and analyses. Resolution and radius are provided in $m$, the latter is the buffer in which data was averaged around each trees. A zero-radius refers to a point extraction from categorical and location specific data. Polygons with radius data were rasterized to a resolution of $5~m$"
caption_ov_short <- "Data descriptions"

@@

Berlin’s open data provided tree inventories including species, age, location, and circumference which was transformed into diameter.
Note that only street trees in urban, not rural areas or within green spaces, were considered here, but individual trees may grow along streets adjacent to green spaces and parks of varying sizes.
Implausible observations, likely from erroneous data entry, were removed.
Additional manual data processing for quality control was done with a bespoke software  `datacleanr` by Hurley *et al*. (submitted), where obvious outliers or clearly interpolated data were removed; the latter was deemed necessary, as several observations in multiple city districts were derived by linear relationships (i.e., straight-line), which do not capture the ontogenic growth dynamics of trees, and leave no variation related to variables other than age.
All of these operations were recorded, and can be viewed and reproduced via the supplementary code.
Lastly, observations with unlikely diameter-age combinations were identified via the residuals of a generalized linear model between diameter and age with a Gamma log-link distribution: if individual residuals exceeded seven times the median absolute deviation of all residuals, they were removed.
The median absolute deviation (MAD) is comparable to the inter-quartile range, yet more robust to outliers:

>>>>

<!-- <!-- lacked spatially-explicit effect size estimates across the urban-rural space and  --> 

@@

It features nearly 700000 individuals covering 94 genera and some 600 species and/or cultivars, listing information on location, stem diameter (at breast height; $DBH$), and stem height, among other variables, for the majority of street and park trees.

@@

caption <- "Data description used for maps/visualizations and analyses. Resolution and radius are in $m$, the latter is the buffer in which data was averaged around each tree. A zero-radius refers to a point extraction from categorical and location specific data. Polygons with radius data were rasterized to a resolution of $5~m$"
caption_ov_short <- "Data descriptions"

@@

Berlin’s open data provided tree inventories including species, age, location, and circumference which was transformed into diameter.
Note that only street trees in urban, not rural areas or within green spaces, were considered here, but individual trees may grow along streets adjacent to green spaces and parks of varying sizes.
Implausible observations, possibly from erroneous data entry, were removed.
Additional manual data processing for quality control was done with a bespoke software  `datacleanr` by Hurley *et al*. (submitted), where obvious outliers or clearly interpolated data were removed; the latter was deemed necessary, as several observations in multiple city districts were derived by linear relationships (i.e., straight-line), which do not capture the ontogenic growth dynamics of trees, and leave no variation related to variables other than age.
All of these operations were recorded, and can be viewed and reproduced via the supplementary code.
Lastly, observations with unlikely diameter-age combinations were identified via the residuals of a generalized linear model between diameter and age with a Gamma log-link distribution: if individual residuals exceeded seven times the median absolute deviation of all residuals, they were removed.
The median absolute deviation (MAD) is comparable to the inter-quartile range, yet more robust to outliers:
